---
title: "plot_examples"
author: "Scot Morrison"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r}

## Load Libraries

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(beeswarm)
library(gghalves)
library(scales)
library(ggthemes)
library(ggdist)

##Set theme 

theme_set(theme_light())

## set color pallet 

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
 
```

#### First up all credit to Patrick Ward for his countless hours of feedback. I doubt I have anything that isn't influenced by him at this point. 

Ok, so lets take a look at how we can create a raincloud plot and a variation that PW showed me using a half box and scatter

First lets create a random data set

```{r}

n = 100
mean = 5
sd = 4
pass_level = 5.5
pop <- tibble(value = rnorm(n, mean, sd))

pop <- pop %>% 
  mutate(result = ifelse(pop$value > pass_level, "pass", "fail")) %>% 
  mutate(post_value = rnorm(n, mean +15, sd)) %>% 
  mutate(sub_id = 1:100) %>% 
  select(sub_id, value, post_value, everything())

pander::pander(head(pop))  ## visualize our data 

```

## Here is the boxplot with jittered points for sample 

```{r fig.width=7,fig.height=8,fig.align='c'}
rate <- pop %>%
  count(result, "pass")

pass_rate <- ((100 * rate[2,3]/(rate[2,3]+rate[1,3])))

plot1 <- pop %>%
  ggplot(aes(x = value, y = factor(0))) +
  geom_boxplot(width = 0.2) +
  geom_jitter(aes(x = value,
                  fill = result),
              position = position_nudge(y = -0.2),
              shape = 21,
              size = 6,
              color = "black",
              alpha = 0.5) +
  geom_vline(aes(xintercept = pass_level), 
             colour = "#E69F00",
                 size = 0.5,
             alpha = 0.5) +
  scale_fill_manual(values = c("pass" = "#56B4E9", "fail" = "#D55E00"))+
  labs(x = "Value",
       y = "Count",
       title = "Example Plot",
       subtitle = "Because it looks cool",
       ) +
  annotate("text", x = -6, y = 1.5, label = paste("Pass Rate = ", round(pass_rate, 1), "% out of", (rate[2,3]+rate[1,3])), hjust = 0) +
  theme_classic() +
      theme(axis.text.y = element_text(),
        axis.title.y = element_text(),
        axis.text.x = element_text(size = 10, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 18),
        legend.position = "bottom")

plot1
  
```

## Same but with raincloud 

```{r}
plot_rain <- pop %>%
      ggplot(aes(x = value, y = factor(0), fill = result)) +
  stat_slab(aes(thickness = after_stat(pdf*n)),
       scale = 0.7) +
  stat_dotsinterval(side = "bottom", 
        scale = 0.9, 
       slab_size = NA)+
  scale_fill_manual(values = c("pass" = "#E69F00", "fail" = "#56B4E9"))+
  labs(x = "Value",
       y = "Count",
       title = "Example Raincloud Plot",
       subtitle = "Because it looks cool",
       ) +
  # annotate("text", x = -6, y = 1.5, label = paste("Pass Rate = ", round(pass_rate, 1), "% out of", (rate[2,3]+rate[1,3])), hjust = 0) +
  theme_classic() +
      theme(axis.text.y = element_text(),
        axis.title.y = element_text(),
        axis.text.x = element_text(size = 10, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 18),
        legend.position = "bottom")
plot_rain
```

## Lets look at change score visualizations that let us still understand the variance in the populatation 

```{r}

## Pivot longer and refactor  
pop_long <- pop %>% 
  pivot_longer(cols = c(value, post_value),
                names_to = "time",
                values_to = "score")  %>% 
                 mutate(time = factor(time, levels = c('value', 'post_value')))

```

```{r}
## Show the change score 



# Define custom labels
custom_labels <- c('Pre', 'Post')

pop_long %>%
  ggplot(aes(x = time, y = score)) +
  geom_line(aes(group = sub_id, color = sub_id),
            size = 0.25) +
  geom_point(aes(group = sub_id,
                 color = sub_id),
             alpha = 0.7,
             size = 2) +
  geom_half_violin(aes(x = time),
                   data = pop_long %>% filter(time == 'post_value'),
                   fill = "#E69F00",
                   color = "#E69F00",
                   side = 'r',
                   alpha = 0.7) +
  geom_half_violin(aes(x = time),
                   data = pop_long %>% filter(time == 'value'),
                   fill = "#009E73",
                   color = "#009E73",
                   side = 'l',
                   alpha = 0.7) +
  scale_x_discrete(labels = custom_labels) +
  labs(x = "Time Period",
       y = "VO2 max",
       title = "VO2 Changes over time") +
  theme(axis.text = element_text(face = "bold", size = 12),
        axis.title = element_text(face = "bold", size = 12),
        plot.title = element_text(size = 18),
        legend.position = "false")
```

#Updated change score visual 

```{r}
# Define a colorblind-friendly palette
cb_palette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Calculate the overall change score
overall_change <- dat_sliced %>%
  group_by(time) %>%
  summarize(mean_score = mean(score, na.rm = TRUE)) %>%
  summarize(change = diff(mean_score)) %>%
  pull(change)

# Calculate individual change scores and join back to the main dataset
dat_sliced <- dat_sliced %>%
  group_by(sub_id) %>%
  mutate(change = score[time == "Post"] - score[time == "Pre"]) %>%
  ungroup()

# Calculate quartiles for each time period
quartiles <- dat_sliced %>%
  group_by(time) %>%
  summarize(q1 = quantile(score, 0.25),
            q2 = quantile(score, 0.5),
            q3 = quantile(score, 0.75))

# Create the plot
ggplot(dat_sliced, aes(x = time, y = score)) +
  geom_half_boxplot(aes(x = time),
                    data = dat_sliced %>% filter(time == 'Pre'),
                    fill = "#56B4E9",
                    color = "black",
                    side = 'l',
                    alpha = 0.7) +
  geom_half_boxplot(aes(x = time),
                    data = dat_sliced %>% filter(time == 'Post'),
                    fill = "#E69F00",
                    color = "black",
                    side = 'r',
                    alpha = 0.7) +
  geom_line(aes(group = sub_id, color = change < 0),
            alpha = 0.8,
            size = 0.85) +
  geom_point(aes(color = time),
             size = 4,
             alpha = 0.7) +
  geom_text_repel(data = dat_sliced %>% 
                    filter(time == "Post"),
                  aes(label = sprintf("%s (%+.1f)", sub_id, change)),
                  size = 3,
                  nudge_x = 0.82,
                  direction = "y",
                  hjust = 0,
                  segment.size = 0.2,
                  segment.color = "gray50") +
  scale_color_manual(values = c("Pre" = "#56B4E9", "Post" = "#E69F00", "TRUE" = "#D55E00", "FALSE" = "gray50")) +
  scale_x_discrete(labels = c("Pre", "Post")) +
  labs(x = "Time Period",
       y = "Bat Speed (mph)",
       title = "FYP 2024 Bat Speed Change Pre to Post", 
       subtitle = paste("Overall Change:", round(overall_change, 2), "mph")) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 20, face = "bold"),
    axis.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 14, face = "bold"),
    axis.text = element_text(color = "black"),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank()
  ) +
  # Add quartile labels
  geom_text(data = quartiles, 
            aes(x = time, y = q1, label = round(q1, 1)),
            vjust = 1.5, hjust = ifelse(quartiles$time == "Pre", 1.2, -0.2), 
            size = 3) +
  geom_text(data = quartiles, 
            aes(x = time, y = q2, label = round(q2, 1)),
            vjust = 1.5, hjust = ifelse(quartiles$time == "Pre", 1.2, -0.2), 
            size = 3) +
  geom_text(data = quartiles, 
            aes(x = time, y = q3, label = round(q3, 1)),
            vjust = -0.5, hjust = ifelse(quartiles$time == "Pre", 1.2, -0.2), 
            size = 3)
```


##trackman
```{r}
## Lets look at change score visualizations that let us still understand the variance in the populatation starting with avg bat speed 




dat_sliced <- tman_dfb %>%
  dplyr::select(batter_name, date, exit_speed, angle) %>%
  rename(name = batter_name) %>%
  arrange(name, date) %>%
  group_by(name, date) %>%
filter(exit_speed >= quantile(exit_speed, 0.9, na.rm = TRUE)) %>%  
  summarize(exit_speed = mean(exit_speed, na.rm = TRUE )) %>%
  mutate(first_date = min(date),
         last_date = max(date)) %>%
  filter(date == first_date | date == last_date) %>%

  mutate(time = factor(ifelse(date == first_date, "Pre", "Post"), levels = c("Pre", "Post"))) %>%
  ungroup() %>%
  select(-first_date, -last_date) %>%
  rename(sub_id = name, score = exit_speed)

# ## Pivot longer and refactor  
# pop_long <- dat %>% 
#   pivot_longer(cols = c(value, post_value),
#                 names_to = "time",
#                 values_to = "score")  %>% 
#                  mutate(time = factor(time, levels = c('value', 'post_value')))

```



```{r}
# Define a colorblind-friendly palette
cb_palette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Calculate the overall change score
overall_change <- dat_sliced %>%
  group_by(time) %>%
  summarize(mean_score = mean(score, na.rm = TRUE)) %>%
  summarize(change = diff(mean_score)) %>%
  pull(change)

# Calculate individual change scores and join back to the main dataset
dat_sliced <- dat_sliced %>%
  group_by(sub_id) %>%
  mutate(change = score[time == "Post"] - score[time == "Pre"]) %>%
  ungroup()

# Calculate quartiles for each time period
quartiles <- dat_sliced %>%
  group_by(time) %>%
  summarize(q1 = quantile(score, 0.25),
            q2 = quantile(score, 0.5),
            q3 = quantile(score, 0.75))

# Create the plot
ggplot(dat_sliced, aes(x = time, y = score)) +
  geom_half_boxplot(aes(x = time),
                    data = dat_sliced %>% filter(time == 'Pre'),
                    fill = "#EF3340",
                    color = "black",
                    side = 'l',
                    width = 0.3,
                    alpha = 0.7) +
  geom_half_boxplot(aes(x = time),
                    data = dat_sliced %>% filter(time == 'Post'),
                    fill = "#00A3E0",
                    color = "black",
                    side = 'r',
                    width = 0.3,
                    alpha = 0.7) +

  geom_line(aes(group = sub_id, color = change < 0),
            alpha = 0.8,
            size = 0.85) +
  geom_point(aes(color = time),
             size = 4,
             alpha = 0.7) +
  geom_text_repel(data = dat_sliced %>% 
                    filter(time == "Post"),
                  aes(label = sprintf("%s (%+.1f)", sub_id, change)),
                  size = 4,
                  nudge_x = 0.82,
                  direction = "y",
                  hjust = 0,
                  segment.size = 0.2,
                  segment.color = "gray50") +
  scale_color_manual(values = c("Pre" = "#EF3340", "Post" = "#00A3E0", "TRUE" = "#D55E00", "FALSE" = "gray50")) +
  scale_x_discrete(labels = c("Pre", "Post")) +
  labs(x = "Time Period",
       y = "90th % Ball Exit Speed (mph)",
       title = "FYP 2024 90th % Exit Speed Change Pre to Post", 
       subtitle = paste("Overall Change:", round(overall_change, 2), "mph")) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 20, face = "bold"),
    axis.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 14, face = "bold"),
    axis.text = element_text(color = "black"),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank()
  ) +
  # Add quartile labels
  geom_text(data = quartiles, 
            aes(x = time, y = q1, label = round(q1, 1)),
            vjust = 1.5, hjust = ifelse(quartiles$time == "Pre", 1.2, -0.2), 
            size = 3) +
  geom_text(data = quartiles, 
            aes(x = time, y = q2, label = round(q2, 1)),
            vjust = 1.5, hjust = ifelse(quartiles$time == "Pre", 1.2, -0.2), 
            size = 3) +
  geom_text(data = quartiles, 
            aes(x = time, y = q3, label = round(q3, 1)),
            vjust = -0.5, hjust = ifelse(quartiles$time == "Pre", 1.2, -0.2), 
            size = 3)

```
